### 2022-03-04 newvariant.tex
done

Analyses of new variants have characterized \emph{relative} strength (i.e., the ratio of reproduction numbers of the invading and resident strains, here called $\rho$) and speed (i.e., the difference between growth rates of the invading and resident strains, here called $\delta$).
As an example, we consider a new variant invading a wild type strain in this paper and use $\Rv$ and $\Rw$ to denote their reproduction numbers and $\rv$ and $\rw$ to denote their growth rates at a given time (and therefore $\rho = \Rv/\Rw$ and $\delta = \rv-\rw$).
Many analyses have focused on changes in the \emph{proportion} of a new variant to estimate its relative speed \citep{switzerland2021variant, davies2021estimated, di2021impact, graham2021changes, leung2021early, volz2021transmission,zhao2021}.
Focusing on proportions can be advantageous, because changes in proportions are less sensitive to changes in testing and to other transient phenomena that would affect variants and wild type viruses
similarly;
however, estimates of relative speed from changes in the proportion of a new variant have typically relied on the assumption that the relative speed remains fixed over the time scale of an invasion.
Instead (or additionally), some studies have assumed a fixed value of the relative strength and tried to predict relative speed \citep{davies2021estimated,di2021impact}.

While both approaches are reasonable, holding different quantities constant (i.e., strength or speed) can lead to different conclusions about the spread of the pathogen and its control \citep{doi:10.1098/rspb.2020.1556}.
To illustrate the differences in conclusions when holding $\RR$ or $r$ fixed, we consider two idealized interventions of constant strength and constant speed.
Given a pre-intervention kernel $K_{\mathrm{pre}}(\tau) = \RR g(\tau)$, which represents the rate at which an infected individual generates secondary infections $\tau$ time units after infection, a constant-strength intervention reduces transmission by a constant factor $\theta$ across generation intervals such that the post-intervention kernel is $K_{\mathrm{post}} = K_{\mathrm{pre}}(\tau)/\theta$.
In this case, the intervention strength $\theta$ must be greater than the epidemic strength $\RR$ to control the epidemic.
In contrast, a constant-speed intervention reduces transmission after infection by a constant rate $\phi$ across generation intervals: $K_{\mathrm{post}} = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$.
In this case, the intervention speed $\phi$ must be greater than the epidemic speed $r$ to control the epidemic \citep{doi:10.1098/rspb.2020.1556}.

@@

Even though SARS-CoV-2 has been spreading for more than a year, there is still considerable uncertainty about its generation intervals.
Several studies have tried to estimate the generation-interval distribution, with means ranging between 3--6 days and squared coefficients of variation ranging between 0.1--1 \citep{ferretti2020quantifying,Ferretti2020timing,ganyani2020estimating,knight2020estimating}.
However, these estimates are derived from serial intervals (i.e., time between symptom onset of the infector and the infectee; \cite{svensson2007note}), which are subject to dynamical biases \citep{park2021forward} and fail to account for asymptomatic transmission \citep{park2020time}.
A recent study has further suggested that the generation intervals of the original SARS-CoV-2 strain may be considerably longer than previously thought \citep{sender2021unmitigated}, adding further uncertainty.
A few studies have tried to compare generation- and serial-interval distributions of new SARS-CoV-2 variants with those of the original wild type strain \citep{pung2021serial, hart2021generation};
these comparisons are inherently difficult due to temporal changes in generation- and serial-interval distributions caused by dynamical and intervention effects \citep{ali2020serial,park2021forward}.
Future studies should prioritize detailed assessment of the generation intervals of SARS-CoV-2 and widespread variants, as well as consider how uncertainty in generational intervals might bias conclusions \citep{doi:10.1098/rsif.2020.0144,ali2020serial,gostic2020practical}.

>>>>

Analyses of new variants have characterized \emph{relative} strength (i.e., the ratio of reproduction numbers of the invading and resident strains, here called $\rho$) and speed (i.e., the difference between growth rates of the invading and resident strains, here called $\delta$).
As an example, we consider a new variant invading a wild type strain in this paper and use $\Rv$ and $\Rw$ to denote their respective reproduction numbers and $\rv$ and $\rw$ to denote their respective growth rates at a given time (and therefore $\rho = \Rv/\Rw$ and $\delta = \rv-\rw$).
Many analyses have focused on changes in the \emph{proportion} of a new variant to estimate its relative speed \citep{switzerland2021variant, davies2021estimated, di2021impact, graham2021changes, leung2021early, volz2021transmission,zhao2021}.
Focusing on proportions can be advantageous, because changes in proportions are less sensitive to changes in testing and to other transient phenomena that would affect variants and wild type viruses
similarly;
however, estimates of relative speed from changes in the proportion of a new variant have typically relied on the assumption that the relative speed remains fixed over the time scale of an invasion.
Instead (or additionally), some studies have assumed a fixed value of the relative strength and tried to predict relative speed \citep{davies2021estimated,di2021impact}.

While both approaches are reasonable, holding different quantities constant (i.e., strength or speed) can lead to different conclusions about the spread of the pathogen and its control \citep{doi:10.1098/rspb.2020.1556}.
To illustrate the differences in conclusions when holding $\RR$ or $r$ fixed, we consider two idealized interventions of constant strength and constant speed.
Given a pre-intervention kernel $K_{\mathrm{pre}}(\tau) = \RR g(\tau)$, which represents the rate at which an infected individual generates secondary infections $\tau$ time units after infection, a constant-strength intervention reduces transmission by a constant factor $\theta$ across generation intervals such that the post-intervention kernel is $K_{\mathrm{post}} = K_{\mathrm{pre}}(\tau)/\theta$.
In this case, the intervention strength $\theta$ must be greater than the epidemic strength $\RR$ to control the epidemic.
In contrast, a constant-speed intervention reduces transmission after infection by a constant rate $\phi$ across generation intervals: $K_{\mathrm{post}} = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$.
In this case, the intervention speed $\phi$ must be greater than the epidemic speed $r$ to control the epidemic.

@@

Even though SARS-CoV-2 has been spreading for more than a year, there is still considerable uncertainty about its generation intervals.
Several studies have tried to estimate the generation-interval distribution, with means ranging between 3--6 days and squared coefficients of variation ranging between 0.1--1 \citep{ferretti2020quantifying,Ferretti2020timing,ganyani2020estimating,knight2020estimating}.
However, these estimates are derived from serial intervals (i.e., time between symptom onset of the infector and the infectee; \cite{svensson2007note}), which are subject to dynamical biases \citep{park2021forward} and fail to account for asymptomatic transmission \citep{park2020time}.
A recent study has further suggested that the generation intervals of the original SARS-CoV-2 strain may be considerably longer than previously thought \citep{sender2021unmitigated}, adding further uncertainty.
A few studies have tried to compare generation- and serial-interval distributions of new SARS-CoV-2 variants with those of the original wild type strain \citep{pung2021serial, hart2021generation};
these comparisons are inherently difficult due to temporal changes in generation- and serial-interval distributions caused by dynamical and intervention effects \citep{ali2020serial,park2021forward}.
Future studies should prioritize detailed assessment of the generation intervals of SARS-CoV-2 and widespread variants, as well as consider how uncertainty in generational intervals might bias conclusions \citep{doi:10.1098/rsif.2020.0144,ali2020serial,gostic2020practical}.
Combining sequencing or viral load trajectory data can also help narrow estimates of generation- and serial-interval distributions as well as time-varying reproduction numbers even when contact tracing data are limited \citep{lin2020viral,Stockdale2022.02.23.22271355}.
